### 2014-12-18 PH.tex
definitions

\pref{def:PH} introduces the Process Hitting~(PH)~\cite{pauleve2011modelisation}
which allows to model
%A Process Hitting (PH) (\pref{def:PH}) gathers 
a finite number of local levels,
called \emph{processes},
grouped into a finite set of components, called \emph{sorts}.
A process is noted $a_i$, where $a$ is the sort's name,
and $i$ is the process identifier within sort $a$.
At any time, exactly one process of each sort is \emph{active},
and the set of active processes is called a \emph{state}.

The concurrent interactions between processes are defined by a set of \emph{actions}.
Actions describe the replacement of a process by another of the same sort
conditioned by the presence of at most one other process in the current state.
An action is denoted by $\PHfrappe{a_i}{b_j}{b_k}$, which is read as
``$a_i$ \emph{hits} $b_j$ to make it \emph{bounce} to $b_k$'',
where $a_i,b_j,b_k$ are processes of sorts $a$ and $b$,
called respectively \emph{hitter}, \emph{target} and
\emph{bounce} of the action.
We also call a \emph{self-hit} any action whose hitter and target sorts are the same,
that is, of the form: $\PHfrappe{a_i}{a_i}{a_k}$.

\begin{definition}[Process Hitting]\label{def:PH} \\
  A \emph{Process Hitting} is a triple $(\PHs,\PHl,\PHa)$:
  \begin{itemize}
    \item[--]  $\PHs = \{a,b,\dots\}$ is the finite set of \emph{sorts};
    \item[--]  $\PHl = \prod_{a\in\PHs} \PHl_a$ is the set of states with
      $\PHl_a = \{a_0,\dots,a_{l_a}\}$
      the finite set of \emph{processes} of sort $a\in\Sigma$
      and $l_a$ a positive integer, with $a\neq b\Rightarrow \PHl_a \cap \PHl_b = \emptyset$;
    \item[--]  $\PHa = \{ \PHfrappe{a_i}{b_j}{b_k} \in \PHl_a \times \PHl_b^2 \mid
      (a,b) \in \PHs^2 \wedge b_j\neq b_k \wedge a=b\Rightarrow a_i=b_j\}$
      is the finite set of \emph{actions}.
  \end{itemize}
\end{definition}

%\begin{definition}[Action]
%\label{def:PhAction}
%An action is noted $\PHfrappe{a_i}{b_j}{b_k}$ where $a_i$ is a process
%of sort $a$ and $b_j$, $b_k$ two processes of sort $b$. When $a_i = b_j$ , such an action is refered as a self-action and $a_i$
%is called a self-hitting process.
%\end{definition}

\begin{example*}
\pref{fig:ph} represents a PH $(\PHs,\PHl,\PHa)$ with three sorts
($\PHs = \{a, b, c\}$) and:
$\PHl_a = \{a_0, a_1\}$,
$\PHl_b = \{b_0, b_1, b_2\}$,
$\PHl_c = \{c_0, c_1\}$.
\begin{figure}[ht]
\centering
\begin{tikzpicture}%[font=\scriptsize]
%\path[use as bounding box] (0,-1) rectangle (4,4);

\TSort{(0,0)}{z}{3}{l}
\TSort{(2,4)}{b}{2}{t}
\TSort{(4,1)}{a}{2}{r}
\THit{b_0}{}{z_1}{.east}{z_2}
\THit{b_1}{}{z_0}{.north east}{z_2}
\THit{a_0}{}{b_1}{.south}{b_0}
\THit{a_1}{out=60,in=0,selfhit}{a_1}{.east}{a_0}

\path[bounce,bend right]
\TBounce{z_1}{}{z_2}{.south}
\TBounce{z_0}{bend right=50}{z_2}{.south east}
;
\path[bounce,bend left]
\TBounce{a_1}{}{a_0}{.north}
\TBounce{b_1}{}{b_0}{.south}
;

 \THit{z_0}{}{a_0}{.west}{a_1}

\path[bounce,bend left]
\TBounce{a_0}{}{a_1}{.south}
;
\TState{a_0,b_0,z_1}
\end{tikzpicture}
\caption{\label{fig:ph} 
A PH model example with three sorts: $a$, $b$ and $z$ ($a$ is either at level 0 or 1, $b$ at either level 0 or 1 and $z$ at either level 0, 1 or 2). Boxes represent the \emph{sorts} (network components), circles represent the \emph{processes} (component levels), and the 4 \emph{actions} that assure the dynamic behavior are drawn by pairs of arrows in solid and dotted lines. The grayed processes stand for a possible initial state.
}
\end{figure}
\end{example*}
A network state is a set of active processes and there is a single process for each kind. In a state $ s \in \PHl$, we say that a process $ a_i \in s$ with $ a $ the name of the sort and $ i $ the level of this sort at this state, and we denote by $s[a]= a_i$.

\begin{definition} [Playable action]
\label{def:playableAction}\\
Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ a state of $PH$. We say that the action $h = \PHfrappe{a_i}{b_j}{b_k} \in \PHa$ is playable at the state $s$ if and only if $hitter(h)=a_i ~ \in s$ and $target(h)=b_j \in s$ (i.e $s[a]=a_i$ et $s[b]=b_j$ ) \\

\end{definition}

The resulting state after playing an action $h$ at an initial state $s$ is denoted by $(s \play h)$ or $(s \play h)[b]=b_k$ and $\forall c \in \PHs, ~ c \neq b, (s \play h)[c]=s[c].$

\subsection*{Dynamic properties}

The study of the dynamics of biological networks was the focus of many researchers, justifying the diversity of network modelisations and the different methods developed in order to verify the dynamic properties. In this paper we present 2 properties: the stable state and the reachability. We develop in this section these properties and how they could be verified in PH network.

The dynamic of a network or its evolution is realised through its actions. In fact when an action $h = \PHfrappe{a_i}{b_j}{b_k}$ is played at a state $s$, the network evolves to the next state results $s'$. We found so in the state $s'$ all processes of $s$ with the replacement of process $b_j=cible(h)$ by $b_k=bond(h)$.

\begin{definition}[Next state]
\label{def:NextState} \\
Let $(\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ be
one of its states. The set of the next possible states for $s$ are computed as follows:
\begin{center}
$next(s) = {s[b_k /b_j ]| \exists (a i , b j ) \in s^2 , \exists b_k \in \PHl_b , \PHfrappe{a_i}{b_j}{b_k} \in \PHa}$
\end{center}
\end{definition}

We call the set of succession of actions ($h_0 \play h_1 \play h_2...$) from a state $ s_0 $ a \emph{scenario} noted $\Sce(s_0)$ corresponding to the succession of states ($s_0 \mapsto s_1 \mapsto s_2 ...$).

\begin{definition}[Stable state or fixed point] \\
\label{def:FixPoint1}
Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
$ \{s_0,..., s_n\} \in \PHl$ be state, $s_n$ is a stable state for $PH$ if and only if $next(s_n) = \emptyset $.
\begin{center}
$s_0 \mapsto s_1 ... \mapsto s_n \mapsto \emptyset $
\end{center}
\end{definition}

\begin{definition}
[Lemme: Stable state or fixed point]
\label{def:FixPoint2}
Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
$s \in \PHl$ be a state, $s$ is a stable state for $PH$ there is no playable action at the state $s$: 
\begin{center}
$\forall h \in \PHa$, $frappeur(h) ~ \notin s ~ \vee ~ cible(h) \notin s$)
\end{center}
\end{definition}

\begin{definition}[Reachability question]
\label{def:reachability}

 If $s \in \PHl$ is a state and $A \in \Proc$ a set of processes, we note $\mathcal{P}(s, A)$ for the \emph{the reachability question}:
 \begin{center}
  $\exists? \delta \in \Sce(s), A \subseteq \PHget(s \play \delta) $ (i.e $\forall a_i \in A, \PHget{(s \play \delta)}{a}=a_i$).
 \end{center}
with $\Sce(s)$ 
is the set of successively playable actions from the state $s$. 
\end{definition}