### 2015-01-15 fixPoint.tex
begin proofread dynamics section

The study of fix points (and, more generally, basins of attraction) provides an important understanding of the different behaviors of a Biological Regulatory Network BRN~\cite{wuensche1998genomic}.
\modMF{%
Indeed, a system will always eventually end in a basin of attraction,
and this may depend on biological switch or other complex phenomena.
}%
A fixed point is a state of the BRN in which it is not possible any more to have new changes;
in other words, it is a basin of attraction that is composed of only one state.

In the following, we consider a Process Hitting $(\Sigma, \PHl, \PHh)$.
It has been shown that a state $s \in L$ is a fixed point of the Process Hitting if and only if $s$ has no next state~\cite{PMR10-TCSB} \ie if there is no playable action in this state.
Therefore, a steady state in a Process Hitting network is a state so that
every process does not hit or is not hit by another process in the same state.
We note that given this result, processes involved in a self-hit cannot be part of a stable state.

\subsection{Process Hitting translation in ASP}
Before writing any piece of ASP,
we first need to present the studied PH network in ASP.
To do this we use the following self-describing predicates:
\texttt{sort}, \texttt{process} and \texttt{action}.
\annotMF{Décrire rapidement le sens de ces prédicats ?}

\begin{example}[Representation of a PH network in ASP]
\label{ex:asp-ph}
The representation of the PH network of Figure \ref{fig:ph} in ASP is the following:
\begin{lstlisting}
sort("a"). sort("b"). sort("z"). %\label{ASPsort}
process("a", 0..1). process("b", 0..1). process("z", 0..2). %\label{ASPprocess}
action("a",0,"b",1,0). action("a",1,"a",1,0). action("b",1,"z",0,2). %\label{actions1}
action("b",0,"z",1,2). action("z",0,"a",0,1). %\label{actions2}
\end{lstlisting}
Line \ref{ASPsort} enumerates every sort of network.
In line \ref{ASPprocess} we create the list of processes corresponding to each sort,
for example the sort \texttt{"z"} has 3 processes numbered from \texttt{0} to \texttt{2};
this specific predicate will in fact expand into the three following predicates:
\texttt{process("z", 0)}, \texttt{process("z", 1)}, \texttt{process("z", 2)}.
Finally, all the actions of the network are defined in lines \ref{actions1} and \ref{actions2};
for example, the first predicate \texttt{action("a",0,"b",1,0)} represents the action
$\PHfrappe{a_1}{b_1}{b_0}$.
\end{example}

\subsection{Search of fix points}
\todo{ Parler du point fixe et ses interrêts. \\
Décrire la nouvelle méthode qui permet de déterminer les états stables d'un réseau }

The enumeration of fix points requires to translate the definition of a stable state
in a set of ASP rules.
The first step consists of eliminating all processes involved in a self-hit;
the others are recorded by the predicate \texttt{shownProcess}:
\begin{lstlisting}
hiddenProcess(A,I) :- action(A,I,B,J,K), A=B, process(A,I),process(B,J),
                      process(B,K). %\label{hiddenProcess}
shownProcess(A,I) :- not hiddenProcess(A,I), process(A,I). %\label{shownProcess}
\end{lstlisting}
Then, we have to browse all remaining processes of this graph
in order to generate all possible states,
that is, all possible combinations of processes by choosing one process from each sort:
\begin{lstlisting}
1 { selectProcess(A,I) : showProcess(A,I) } 1 :- sort(A).
\end{lstlisting}
\modMF{%
The previous line creates an many candidate answer sets as there are possible states
to take into account.
The last step consists of filtering each state that is not a fixed points,
or, in other words, eliminating all candidate answer sets that do not
represent a fixed point.
For this, we use a constraint:
indeed, any solution that satisfy the body of a constraint will be removed from the answer set.
}%
Regarding our problem, a state is eliminated if there exists an action linking two processes:
\begin{lstlisting}
:- hit(A,I,B,J), selectProcess(A, I), selectProcess(B, J), A != B. %\label{contraintFix}
\end{lstlisting}

\begin{example}
The PH model of figure \pref{fig:ph} contains 3 sorts:
$a$ and $b$ have 2 processes and $z$ has 3, therefore, the whole model has $2*2*3 = 12$ states (whether they can be reached or not from a given initial state).
We can check that this model contains only one fixed point: $\PHstate{b_0, z_2, a_0}$.
Indeed, there is no action between each two of the processes contained is this state;
moreover, no other state verifies this property.

If we execute the ASP program detailed below,
alongside with the description of the PS model given in example~\ref{ex:asp-ph},
we obtain one answer set, which matches the expected result:
\begin{lstlisting}[numbers=none]
Answer 1 : fixProcess(a, 0), fixProcess(b, 0), fixProcess(z, 2)
\end{lstlisting}
\end{example}