### 2015-01-28 comparaison.tex

In this section, we show the effectiveness of our approach on some examples,
and compare it to other existing approaches.
All computations were performed on a Pentium~V, 3.2~GHz with 4~GB RAM.

\subsection{Evaluation}
To evaluate the effectiveness of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method for CTL model-checking of Rocca's et al. \cite{roccaasp} which was developed also on ASP but for transition states networks.
Each method use a specific kind of representation\footnote{The transalation from PH to other models is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
Instantiable Transition Systems for \textsc{libDDD} is a library for symbolic model-checking of CTL \& LTL properties. \footnote{\textsc{libDDD} is \url{http://www.lip6.fr/} },
state transitions network for Rocca's et al. method
and Process Hitting (PH) for \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models. It available online \url{http://loicpauleve.name/pint/} } as well as for our method.

For this comparative study, we focus biological network of different sizes:
a tadpole tail resorption (TTR) model with 12 biological components~\cite{khalis2009smbionet},
an ERBB receptor-regulated G1/S transition (ERBB) model with 20 components~\cite{Samaga2009}
and a T-cell receptor (TCR) signaling network of 40 components~\cite{Klamt06}.

These models were chosen to be of different sizes:
from small (12 components) to large (40 components).
We note however that the PH models considered may contain more sorts than
the original number of biological components, due to the use of
“cooperative sorts”, which allow to model Booelan gates but do not necessarily
have a biological meaning.
The different model representations required have been obtained by translations
from the PH
ensuring the conservation of the dynamical properties.
All these translations are available with \textsc{PINT} 
All results alongside with more detailed specifications of the models
are given in table \ref{tab:reachability}.
The methods and the results provided by each of them are detailed in the following.
The overall results shows that our method is efficient in computing reachability
from a given initial state;
furthermore, it sometimes provides more informations that the other existing ones.

\begin{figure}[htp]
\begin{center}
\label{tab:reachability}
\noindent%
\begin{tabular}{|l|c|c|c|c||c|c|c|c|c|}
  \hline
   Model&  \#sorts &\#procs & \#states & target & \textsc{ASP-CTL} & \textsc{Pint} & \textsc{libddd} & \textsc{GINsim} & \textsc{ASP-PH} \\
  \hline
  TTR & 12 &42 & $2^{19}$ & full state & 7m17.01s & - & 0m1.151s &  0m1.001s & 0m1.90s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & full state & - & - &1m55.38s & 2m31.64s & 0m11.84s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & partial state & - & 0m0.027s &1m54.96s & - & 0m5.02s \\
  \hline
  TCR & 40 &156 & $2^{73}$ & full state & - & - & out & out & 4m27.93s \\
  \hline
   TCR & 40 &156 & $2^{73}$ & partial state & - & 0m0.014s & out & - & 1m35.080s \\
  \hline
\end{tabular}
\caption{Compared performances of Rocca et al. method denoted by \textsc{ASP-CTL}, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method \textsc{ASP-PH} developped in ASP for Process Hitting networks.
For each test, this table gives the considered model,
the number of its processes and its states, the type of goal
(either a whole state or a sub-state)
and the computation time of the different methods used for the tests,
where “out” marks an execution taking too much time or memory
and “-” indicates that is not possible to do the test.
}
\end{center}
\end{figure}

\todo{Diviser en 2 tableaux}

\todo{Il n'y a pas de résultats sur le calcul des poitns fixes dans le tableau} \repEB{En fait pour la méthode de Rocca le script qu'il m'ont envoyé génére quelques erreurs lors de la compilation. pour libddd y'avait pas l'option (--fixpoint) ne génère rien comme résultat et GINsim répond très rapidement mais il faudrait calculer tout le graphe de transition avant. }

\begin{itemize}

\item \textbf{Rocca's method} offers the possibility to model-check CTL properties
of state-transition networks.
We note that in general, computing such a graph from a network is exponential in the number of components and local states in the model.
Therefore, using this method first requires an exponential computation to
express the state-transition graph of the studied model into ASP.
Indeed, for the tadpole tail resorption model comprising only 42 processes,
this computation took 3 minutes, which was not included in the solving time
reported in table~\ref{tab:reachability}.
Afterwards, this method is able to check any kind of CTL formula
(and not only the EF form that we focused on in this paper),
but this checking may be time-consuming.

\item \textbf{\textsc{GINsim}} is a software for the edition, simulation and analysis
of gene interaction networks.
It allows to compute all reachable stable states instantly;
however, it is not possible to compute all stable states independently from the initial state.
Regarding the reachability problem, \textsc{GINSim} only allows to check
full states, because of its approach which consists in computing
all the state-transition graph and then check if there exists a path between two given states.
Therefore, it was not possible to perform reachability checks on partial states.
Small state-transition graphs can also be displayed by this tool.

\item \textbf{\textsc{libddd}}
is a library for symbolic model-checking of CTL \& LTL properties.
It can thus especially be used to check reachability properties;
however, as opposed to our method, it does not output an execution path
solving this reachability.
In addition, it relies on the construction of the state-transition graph
which is then stored under the form of a binary decision diagram for a more efficient analysis.

This computation explains why \textsc{libddd} takes more time to respond,
and does not answer in less than 12 minutes for the biggest example
(which contains $2^{73}$ states).
Finally, \textsc{libddd} is not able to compute the stable states of a network.

\item \textbf{PINT}
is a library gathering tools and translators related to the PH.
It should be noted that Pint contains the only reachability analysis
developed so far natively for the Process Hitting,
before the method proposed in this paper.
It consists in an approximation that avoids to compute the state-transition graph;
It is thus ensured to be really efficient, which explains the fastest results,
but at the cost of possibly terminating without being conclusive (although this is rare).
However, it is not designed for goals consisting of a full state, which
may trigger an exponential process to reach an answer,
which explains the high computing times for some of them \ref{tab:reachability}
Moreover, in the case of a positive answer,
it currently does not return the execution the path achieving the desired reachability,
but only outputs its conclusion.
\end{itemize}

\subsection{Strengths and limitations of our method}

In the previous sections,
we developed new methods to check dynamical properties,
namely identifying stable states and finding all possible paths to reach a given goal.
Compared to other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}),
our method is relatively faster and also permits to study larger networks
(up to $2^{73}$ states in our tests) despite some inconclusive cases.
Indeed, our ASP program is non-conclusive in the cases where the given goal is not
reachable and the model contains loops in its dynamics
(which happens in almost every biological model).
In this case, the program will compute all infinite paths in these loops,
and never reach a goal or a fixed point.
It is still possible, however, to limit the number of iterations to an arbitrary
maximum which will be eventually reached in the case of an endless loop.
This is possible with the option \texttt{"-{}-imax=n"} of \textsc{iclingo},
where \texttt{n} is the maximum number of steps.
%%
%\begin{lstlisting}[numbers=none]
%iclingo find\_path.lp <network\_name>.lp -- imax=n
%\end{lstlisting}
%%
For example, the total number of states is an obvious maximum,
as it will never be exceeded by a minimum path,
but it is too hight to be very interesting.
The total number of sorts is a more interesting value,
under the hypothesis that each one will change its active process at most once,
which is often the case for Boolean networks;
or, with a similar reasoning, the total number of processes can be chosen.
In this case, however, a termination with no solution cannot be considered as a formal
negative answer, unless one can prove that the chosen value \texttt{n}
is bigger than the longest possible path in the state-transition graph.
Given our implementation, if the step \texttt{n} is reached,
the computation stops and displays paths that are not valid executions
to the goal.
\todo{Y a-t-il moyen de les différencier de chemins corrects?}