### 2015-04-20 main.tex
changement pour ICLP-15

% TLP2esam.tex / sample pages for TLP
% v2.11, released 6-nov-2002

\documentclass{tlp}
%\usepackage{aopmath}

% Packages
\usepackage{latexsym}
\usepackage{epic,eepic}
\usepackage{times}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multirow}
\usepackage[dvips]{graphicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{xcolor}
\usepackage{ulem}
\usepackage{amssymb}

%Package CAV
\usepackage[english]{babel}
\usepackage{stmaryrd} % Maths (crochets doubles)
\usepackage{url}     % Mise en forme + liens pour URLs
%\usepackage{array}   % Tableaux évolués
\usepackage{setspace}

% Tunning
%------------------
\usepackage[12pt]{moresize}
\algtext*{EndIf}% Remove "end if" text
\algtext*{EndFor}% Remove "end if" text
%\setlength{\belowcaptionskip}{10.0pt}
%\setlength{\abovecaptionskip}{10.0pt}
%\setlength{\textfloatsep}{10.0pt}
%\setlength{\floatsep}{10.0pt}
% TUNNING
\setstretch{0.94}
%\fontsize{3.85mm}{3.85mm}\selectfont

\newcommand{\pf}[1]{\langle#1\rangle}
\newcommand{\noproof}{\hfill $qed$}
\newcommand{\ignore}[1]{}

\definecolor{gray50}{gray}{0.15}
\graphicspath{{pictures/}}

\newtheorem{definition}{Definition} % [section]
\newtheorem{example}{Example} % [section]
\newcommand{\pivot}[1]{\mathbin{\, {#1} \,}}
\newcommand{\Pivot}[1]{\mathbin{\; {#1} \;}}
\let\from=\leftarrow

%Commande perso
\newcommand{\ie}{i.e.,\ }
\newcommand{\eg}{e.g.,\ }
\newcommand{\resp}{resp.\ }

\usepackage{listings}
% Définition du langage ASP
\lstdefinelanguage{ASP}{%
literate={:-}{$\leftarrow$}1,
morekeywords={not},
numbers=left,
escapeinside={\%}{\^^M}
}
% Définition des styles de tous les listings du document
\lstset{
language=ASP,
basicstyle=\small\ttfamily,
columns=fullflexible,
keywordstyle=\bfseries,
firstnumber=last,
keepspaces=true
}
\renewcommand{\thelstnumber}{\the\value{lstnumber}}
%% fin définition

% Styles ASP inline
\newcommand{\ASPnot}{\textbf{not}~}

% les inputs files
\input{macros/macros}
\input{macros/macros-ph}
\input{macros/tikzstyles2.tex}
%\input{macros/macros-abstr}

\begin{document}
\bibliographystyle{acmtrans}

\long\def\comment#1{}

\title[Exhaustive analysis of biological regulatory networks using ASP]{Exhaustive dynamic analysis of reachability and stable states
of biological regulatory networks modeled in Process Hitting
using Answer Set Programming}

\author[E. Ben Abdallah, M. Folschette, O. Roux, M. Magnin]
{Emna Ben Abdallah\\
$^1$LUNAM Université, \'Ecole Centrale de Nantes, \\
 IRCCyN UMR CNRS 6597\\ (Institut de Recherche en Communications et Cybern\'etique de Nantes),\\
  1 rue de la Noë, 44321 Nantes, France.\\
\email{emna.ben-abdallah@irccyn.ec-nantes.fr},\\
%E-mail: 
\and Maxime Folschette$^{1,2}$\\
$^2$School of Electrical Engineering and Computer Science,\\
University of Kassel, Germany \\
\email{maxime.folschette@uni-kassel.de},\\
\and
Olivier Roux$^1$ \\
LUNAM Université, \'Ecole Centrale de Nantes, IRCCyN \\
\email{olivier.roux@irccyn.ec-nantes.fr},\\
\and Morgan Magnin$^{1,3}$ \\
$^3$National Institute of Informatics, \\
2-1-2, Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan.\\
\email{morgan.magnin@irccyn.ec-nantes.fr},\\
}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\volume{\textbf{10} (3):}
\jdate{April 2015}
\setcounter{page}{1}
\pubyear{2015}

\maketitle

\label{firstpage}
%
\begin{abstract}
In this paper, we propose a new logical approach to perform model-checking on a restriction of synchronous automata networks, namely Process Hitting. The Process Hitting framework is notably suitable, but not limited, to model and analyze efficiently biological regulatory networks. 
The originality of our model-checking approach relies in the use of Answer Set Programming to consider exhaustive enumeration of all possible solutions of a property with respect to a given execution length.  Our work here focuses on state reachability properties (which are equivalent to the CTL “$\mathsf{EF}$” operator with paths limited to a given length) on the one hand, on the identification of fixed points on the other hand.
The merits of our methods is illustrated by applying them to
biological examples of various sizes
and comparing the results with some existing approaches.
It turns out that our approach succeeds in processing large models
with a high number of components and interactions.

	\keywords{Process Hitting, Answer Set Programming, stable states, fixed points, reachability}
\end{abstract}

\section{Introduction}
\input{parts/intro}

\section{Preliminary definitions}
\label{defs}
%\input{ASP}
\subsection{Process Hitting}
\input{parts/PH}

\section{Fixed point enumeration}
\label{fixpoint}
\input{parts/fixPoint}

\section{Dynamical analysis}
\label{dynamics}
\input{parts/reachability}

\section{Comparative performance analysis}
\label{comparison}
\input{parts/comparaison}

%\vspace{-1.0cm}

\section{Conclusion and future directions}
\label{ccl}
\input{parts/cc}

%\clearpage
\bibliographystyle{acmtrans}
\bibliography{biblio}

\clearpage
\appendix

%\ssmall
\input{parts/ASP}

\end{document}